Query Optimization for Dynamic Imputation

### 2017-05-15 paper/examples.tex

For the query in~\Cref{fig:example-query}, \ProjectName{} generates
the plans in~\Cref{fig:quality-plan} and~\Cref{fig:fast-plan}, when 
optimizing for quality and performance, respectively.

\Cref{fig:quality-plan} shows a quality-optimized plan that uses the impute operator $\mu$,
which employs a reference imputation strategy to fill in missing values rather than dropping tuples.
It waits to impute \verb|demo.income| until after the final join has taken place, though other imputations take place earlier on in the plan, some before filtering and join operations.
Imputations are placed to maximize ImputeDB's estimate of the quality of the overall results.
Meanwhile, \Cref{fig:fast-plan}'s performance-optimized plan uses the drop operation $\delta$ instead of filling in missing values.
However, it is a significant improvement over dropping all tuples with missing data in any field, as it only drops tuples with missing values in attributes which are referenced by the rest of the plan. In both cases, only performing
imputation on the necessary data yields a query that is much faster
than performing imputation on the whole dataset and then executing
the query.

@@

Tens of queries later, our epidemiologist has a holistic view of the data and she has the information that she needs to construct a tailored imputation model for her queries of interest.
Knowing that there may be a need to explore this data set further, she can also incorporate her imputation model into ImputeDB for future use.

>>>>

For the query in~\Cref{fig:example-query}, \ProjectName{} generates
several plans on an optimal frontier. \Cref{fig:quality-plan} shows a quality-optimized plan that uses the impute operator $\mu$,
which employs a reference imputation strategy to fill in missing values rather than dropping tuples.
It waits to impute \verb|demo.income| until after the final join has taken place, though other imputations take place earlier on in the plan, some before filtering and join operations.
Imputations are placed to maximize ImputeDB's estimate of the quality of the overall results.
On the other hand, \Cref{fig:fast-plan} shows a performance-optimized plan that uses the drop operation $\delta$ instead of filling in missing values.
However, it is a significant improvement over dropping all tuples with missing data in any field, as it only drops tuples with missing values in attributes which are referenced by the rest of the plan. In both cases, only performing
imputation on the necessary data yields a query that is much faster
than performing imputation on the whole dataset and then executing
the query.

@@

Tens of queries later, our epidemiologist has a holistic view of the data and she has the information that she needs to construct a tailored imputation model for her queries of interest.
